---
title: "lab-07-simpsons.Rmd"
author: "nawal faleh almutairi "
date: "17 March 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r load_libraries, message=FALSE}
library(tidyverse) 
library(mosaicData)
```


## Exercises


1.  
```{r whickam}
?Whickham
```
    
Your answer:
the data is observational as the description states that is based on age, smoking, and mortaliy, which are all observable events and not produced via experiments.  

2.  
```{r observations}
nrow(Whickham)
```

Your answer;there are 1,314 observations. as we know every row is an observation

3.  

```{r number_variables}
names(Whickham)
```
Your answer:
Your answer:there are 3 variables, "outcome", "smoker", and "age"

```{r var_type}
unique(Whickham$outcome)
unique(Whickham$smoker)
unique(Whickham$age)
```

Your answer: 
Using the unique() function on the 3 variables we could see that "outcome" only takes Alive or Dead value, which makes it categorical non-ordinal. "smoker" only takes yes or no, which also makes it categorical non-ordinal. Age is numerical continous data.                                                                                                                
one of the best ways to visualis categorical data is through the use of bar charts
```{r plot_outcome}
ggplot(Whickham, aes(x = outcome)) +
 geom_bar()
```


```{r plot_smoker}
ggplot(data=Whickham, aes(x=smoker, y=outcome, color=outcome)) + geom_bar(stat="identity")
```

```{r plot_age}
ggplot (Whickham, aes(x = age)) +geom_boxplot()
```
ANSWER4
4. One out of every two smokers dies of a disease caused by smoking cigarettes, directly.

```{r plot_smokeVSincome_1}
ggplot(data=Whickham, aes(x=smoker, y=outcome, color=outcome)) + geom_bar(stat="identity")    
```

Knit, commit, and push to github.

5. Answer5 
Tobacco kills nearly half of those who use it.
Globally, about 80% of smokers live in low and middle income countries.
Tobacco contains tar, which causes lung cancer and narrowing of the airways.
Reduces the efficiency of red blood cells in transporting oxygen due to its carbon monoxide content; Which leads to:
Arteriosclerosis, high blood pressure.
Emphysema and bronchitis, Heart attack (angina),Optic atrophy.
brain attack, Teeth falling out, Cardiac infarction.

```{r plot_smokeVSincome_2}
Whickham %>%
  count(smoker, outcome)
502+230
 230/732
```

6.

```{r age_cat}
Whickham <- Whickham%>% mutate (age_cat = case_when (age <= 44 ~ "18-44", age > 44. & age <= 64 ~ "45-64", age > 64 ~ "65+"))
```

7. 

```{r plot_smokeVSincome_3}
ggplot(data=Whickham, aes(x=outcome, y=smoker,color=smoker, fill=age_cat)) + geom_bar(stat="identity") +facet_wrap(~age_cat)
```

Knit, commit, and push to github.